A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-12 manuscript.Rmd
clarified checksums

```{bash, eval=FALSE}
choco install -y git make docker-desktop
```

@@

```{bash, eval=FALSE}
brew cask docker
```

@@

```{bash, eval=FALSE}
apt install git make docker 
```

@@

Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts. In most of the cases we have seen, these are either space constraints or privacy considerations. In these cases, of course, unrestricted reproducibility is not guaranteed. If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them. To do this, a checksum must be created and stored at the time of the original analysis. At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.

```{r}
x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
checksum <- digest::digest(x, "sha1")
if (checksum != "8e222b9b3616595dc3f5946f9eac101fbf1336dc"){
  warning("Mismatch between original and current data file!")
}
```

>>>>

```{bash, eval=FALSE}
# install docker using choco
choco install -y git make docker-desktop
```

@@

```{bash, eval=FALSE}
# install docker via brew
brew cask docker
```

@@

```{bash, eval=FALSE}
# install docker via advanced package tool
apt install git make docker 
```

@@

Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts. In most of the cases we have seen, these are either space constraints or privacy considerations. In these cases, unrestricted reproducibility is not guaranteed. If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them. A checksum is a short fixed-length fingerprint (often displayed in hexadecimal system) of a file with the purpose of verifying integrity of a digital object. Fingerprints are computed from digital objects such that they change with high probability if data is changed only little. To use checksum validation, checksums for all data files must be created and stored at the time of the original analysis. At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.

```{r}
# create a dummy data.frame with two columns
x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
# compute checksum using md5
checksum <- digest::digest(x, "md5")
if (checksum != "5ba412f5a26f43842971dd74954fcdeb"){
  warning("Mismatch between original and current data file!")
}
```
